Online Supplement to "Demographic costs of sexually antagonistic selection in partially selfing populations \\ \textit{The~American~Naturalist}

### 2021-05-25 doc/SA-Hermaphrodites-wDemography.tex
Code for sex-specific dominance simulations
This commit adds functionality to several simulation and plotting
functions to explore the consequences of sex-specific dominance

\documentclass[11pt,draft]{article}
% Preamble
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage[authoryear,sectionbib,sort]{natbib}
\bibliographystyle{amnatnat.bst}
\setlength{\bibsep}{0.0pt}
\linespread{1.7}
\usepackage[utf8]{inputenc}
\usepackage[left]{lineno}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{color,soul}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pdflscape}
\usepackage{nameref}
%\usepackage[title,titletoc,toc]{appendix}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{pdflscape}

@@

% Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for restricted conditions (see \hl{Online Appendix X}). We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space. 
Because single-locus selection coefficients are generally weak (e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise. For simplicity, we present results for scenarios of equal dominance in the main text (i.e., $h_f = h_m = h$). Specifically, we examine scenarios of additive SA fitness effects ($h = 1/2$) and dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$). \textcolor{red}{We relax these assumptions to explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016}, on our model predictions in \hl{Appendix X}). }

@@

We used flower number as a proxy for female fertility. To calculate selection on flower number, we first calculated the geometric mean flower number in the field for wild-type and inv6 heterozygotes across the two years for which data was available (2012 and 2013), extrapolated the expected flower number for inv6 homozygotes, and calculated the relative fertility of each genotype given that inv6 homozygotes have the highest flower number ($w_i$, where $i \in \{inv6/inv6,\, inv6/w+,\, w+/w+\}$). Using the relative fertility expressions outlined in Table \ref{tab:inv6Fitness}, we then solved for $s_f$.

We used pollen production as a proxy for male fertility. We calculated average pollen production as the product of the geometric mean flower number and field estimated pollen viability for wild-type and inv6 heterozygotes, and extrapolated the expected pollen production for inv6 homozygotes. As before, we calculated relative male fertility ($w^{\prime}_i$) for each genotype given that wild-type homozygotes had the highest pollen viability, and used the expressions for $w^{\prime}_{i}$ from Table \ref{tab:inv6Fitness} to solve for $s_m$.

It is worth noting that \citet{LeeKelly2015} also report greenhouse pollen viabilty data for several mapping populations of {\itshape M.~guttatus}, which includes data for inv6 homozygotes, from which relative fertilities can be estimated. We chose to use the field estimated pollen viability and flower number data, despite the lack of inv6 homozygotes, because it probably gives a more accurate reflection of the fitness effects of inv6 in the field. Moreover, both approaches require making assumptions regarding the dominance effects of inv6. Given the obvious pitfalls of estimating selection on inv6 from the available data, it goes without saying that the location of inv6 in demographically viable polymorphic parameter space is a highly speculative, but nevertheless interesting, proof of concept for our model predictions.

\begin{table}[htbp]
 \centering
 \caption{\bf Relative fertilities for inv6 genotypes ($w_{i}$)}
\label{tab:inv6Fitness}
\begin{tabular}{lccc}
 \toprule
					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
\cline{2-4}
									& $inv6/inv6$	& $inv6/w+$		& $w+/w+$ 	\\ \hline
Female function ($w_{i}$):			& $1$			& $1 - s_f/2$	& $1 - s_f$ \\	
Male function ($w^{\prime}_{i}$):	& $1 - s_m$		& $1 - s_m/2$	& $1$ 		\\	
\hline
\end{tabular}
\end{table}

>>>>

\documentclass[11pt]{article}
% Preamble
\usepackage[sc]{mathpazo} %Like Palatino with extensive math support
\usepackage{fullpage}
\usepackage[authoryear,sectionbib,sort]{natbib}
\bibliographystyle{amnatnat.bst}
\setlength{\bibsep}{0.0pt}
\linespread{1.7}
\usepackage[utf8]{inputenc}
\usepackage[left]{lineno}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{color,soul}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pdflscape}
\usepackage{nameref}
%\usepackage[title,titletoc,toc]{appendix}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\usepackage{pdflscape}

@@

% Unfortunately, closed-form solutions for the invasion conditions for each allele and the equilibrium frequencies under different selection scenarios are analytically intractable except for restricted conditions (see \hl{Online Appendix X}). We therefore used numerical techniques to identify invasion and extinction thresholds, which could then be used to define regions of demographically viable polymorphism across $s_f \times s_m$ parameter space. 
Because single-locus selection coefficients are generally weak (e.g., \citealt{Eyre-WalkerKeightly2007}) and strongly skewed, we limit our analyses to coefficients within $0 < s_f,s_m \leq 0.15$, unless stated otherwise. For simplicity, we present results for scenarios of equal dominance in the main text (i.e., $h_f = h_m = h$). Specifically, we examine scenarios of ({\itshape i}) additive SA fitness effects ($h = 1/2$); and ({\itshape ii}) dominance reversals, where the deleterious fitness effect of each SA allele is partially recessive in each sex ($h = 1/4$). These scenarios are of particular interest because additive effects are commonly observed for small-effect alleles or quantitative traits \citep{AgrawalWhitlock2011}, while dominance reversals are predicted under fitness landscape models provided the population is not too far from the phenotypic optimum \citep{Manna2011, ConnallonClark2014}. \textcolor{red}{We explore the consequences of sex-specific dominance, which may evolve under asymmetric SA selection \citep{SpencerPriest2016} in \hl{Appendix X}).}

@@

We used flower number as a proxy for female fertility. To calculate selection on flower number, we first calculated the geometric mean flower number in the field for wild-type and inv6 heterozygotes across the two years for which data was available (2012 and 2013), extrapolated the expected flower number for inv6 homozygotes (table \ref{tab:MimReproEst}), and calculated the relative fertility of each genotype given that inv6 homozygotes have the highest flower number ($w_i$, where $i \in \{inv6/inv6,\, inv6/w+,\, w+/w+\}$). Using the relative fertility expressions outlined in Table \ref{tab:inv6Fitness}, we then solved for $s_f$.

\begin{table}[htbp]
 \centering
 \caption{\bf Selection coefficient estimates for inv6}
\label{tab:MimReproEst}
\begin{tabular}{lccc}
 \toprule
 	&  \multicolumn{3}{c}{{Genotype}} \\ 
\cline{2-4}
Geom.~Mean (2012 \& 2013)			& $inv6/inv6$	& $inv6/w+$		& $w+/w+$ \\ \hline
Flower \#:								& $4.63$	& $3.91$	& $3.19$ \\	
Flower \# $\times$ Pollen viab.:		& $1.89$	& $2.29$	& $2.70$ \\ \hline
Rel.~Flower \#:							& $1$		& $0.85$	& $0.69$ \\	
Rel.~Flower \# $\times$ Pollen viab.:	& $0.70$	& $0.85$	& $1$ \\	
\hline
\end{tabular}
\caption{\footnotesize{Note that Values for $inv6/inv6$ homozygotes are extrapolated based on the assumption of additive effects of the inversion on flower and pollen production.}}
\end{table}

We used pollen production as a proxy for male fertility. We calculated average pollen production as the product of the geometric mean flower number and field estimated pollen viability for wild-type and inv6 heterozygotes, and extrapolated the expected pollen production for inv6 homozygotes (table \ref{tab:MimReproEst}). As before, we calculated relative male fertility ($w^{\prime}_i$) for each genotype given that wild-type homozygotes had the highest pollen viability, and used the expressions for $w^{\prime}_{i}$ from Table \ref{tab:inv6Fitness} to solve for $s_m$.

\begin{table}[htbp]
 \centering
 \caption{\bf Relative fertilities for inv6 genotypes ($w_{i}$)}
\label{tab:inv6Fitness}
\begin{tabular}{lccc}
 \toprule
					&  \multicolumn{3}{c}{{Genotype}} \\ 
\cline{2-4}
									& $inv6/inv6$	& $inv6/w+$		& $w+/w+$ 	\\ \hline
Female function ($w_{i}$):			& $1$			& $1 - s_f/2$	& $1 - s_f$ \\	
Male function ($w^{\prime}_{i}$):	& $1 - s_m$		& $1 - s_m/2$	& $1$ 		\\	
\hline
\end{tabular}
\end{table}

\citet{LeeKelly2015} also report greenhouse pollen viabilty data for several mapping populations of {\itshape M.~guttatus}, which includes data for inv6 homozygotes, from which relative fertilities can be estimated. We chose to use the field estimated pollen viability and flower number data, despite the lack of inv6 homozygotes, because it probably gives a more accurate reflection of the fitness effects of inv6 in the field. Moreover, both approaches require making assumptions regarding the dominance effects of inv6. Given the obvious pitfalls of estimating selection on inv6 from the available data, it goes without saying that the location of inv6 in demographically viable polymorphic parameter space is a highly speculative, but nevertheless interesting, proof of concept for our model predictions.

### 2021-05-25 doc/Supplements/SupplementaryMaterial.tex
Code for sex-specific dominance simulations
This commit adds functionality to several simulation and plotting
functions to explore the consequences of sex-specific dominance

\bibliographystyle{amnatnat.bst}
\setlength{\bibsep}{0.0pt}
\linespread{1.7}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]
\usepackage{fancyhdr}
\usepackage{color,soul}
\pagestyle{fancy}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\setlength{\headheight}{13.6pt}
\renewcommand{\refname}{Literature Cited}

@@

%%%%%%%%%%%%%%%%%%%%%
% Header
%%%%%%%%%%%%%%%%%%%%%
%
% Customize the line below with the last name of your first author and
% the short title of your MS. You can comment authorship information out
% while your MS is undergoing double-blind review.
%
\rhead{Supplement to Olito \& de Vries., "Costs of SA selection" \textit{Am.~Nat.}}
\setlength{\headsep}{0.3in}  
\lhead{}

@@

\title{Online Supplement: Demographic costs of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites \\ 
%\LaTeX{} Template for Author-Supplied Supplementary PDFs, \\ 
\textit{The~American~Naturalist} }

@@

%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%
%
% Any references in this section should also be copied into the main 
% template, under the heading
% \section*{References Cited Only in the Online Enhancements}
% except when they are already cited in the main text.
\bibliography{Refs2.bib}

>>>>

\bibliographystyle{amnatnat.bst}
\setlength{\bibsep}{0.0pt}
\linespread{1.5}
\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{titlesec}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\titleformat{\section}[block]{\Large\bfseries\filcenter}{\thesection}{1em}{}
\titleformat{\subsection}[block]{\Large\itshape\filcenter}{\thesubsection}{1em}{}
\titleformat{\subsubsection}[block]{\large\itshape}{\thesubsubsection}{1em}{}
\titleformat{\paragraph}[runin]{\itshape}{\theparagraph}{1em}{}[. ]
\usepackage{fancyhdr}
\usepackage{color,soul}
\pagestyle{fancy}
\usepackage[colorlinks=true, allcolors=black]{hyperref}
\setlength{\headheight}{13.6pt}
\renewcommand{\refname}{References}

@@

%%%%%%%%%%%%%%%%%%%%%
% Header
%%%%%%%%%%%%%%%%%%%%%
%
% Customize the line below with the last name of your first author and
% the short title of your MS. You can comment authorship information out
% while your MS is undergoing double-blind review.
%
\rhead{Supplement to Olito \& de Vries, "Costs of SA selection" \textit{Am.~Nat.}}
\setlength{\headsep}{0.3in}  
\lhead{}

@@

\title{Online Supplement to "Demographic costs of sexually antagonistic selection in partially selfing populations \\ 
%\LaTeX{} Template for Author-Supplied Supplementary PDFs, \\ 
\textit{The~American~Naturalist} }

@@

%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%
%
% Any references in this section should also be copied into the main 
% template, under the heading
% \section*{References Cited Only in the Online Enhancements}
% except when they are already cited in the main text.
\bibliography{suppRefs.bib}
